Stochastic Variance-Reduced Policy Gradient

### 2018-02-09 svrpg.tex
Fixes to experiments and conclusions

\section{Experiments}\label{sec:exp}
In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer~\citep[\eg][]{duan2016benchmarking}.
We consider G(PO)MDP since has a smaller variance than REINFORCE.
% This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases. 
For our algorithm, we use a batch size $N=100$, a mini-batch size $B=10$, and the jointly adaptive step size $\alpha$ and epoch length $m$ proposed in Section \ref{sec:stopping}. Since the aim of this comparison is to show the improvement that SVRG-flavored variance reduction brings to SG in the policy gradient framework, we set the batch size of the baseline policy gradient algorithm to $B$. In this sense, we measure the improvement yielded by computing snapshot gradients and using them to adjust parameter updates. Since we evaluate on-line performance over the number of sampled trajectories, the cost of computing such snapshot gradients is automatically taken into consideration. To make the comparison fair, we also use Adam in the baseline PG algorithm, which we will denote simply as G(PO)MDP in the following.
In all the experiments we used deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}).
Each experiment is run $10$ times with a random policy initialization and seed but this initialization is shared among the algorithms under comparison.
The length of the experiment, \ie the total number of trajectories, is fixed for each task. Performance is evaluated by using test-trajectories on a subset of the policies considered during the learning process. We provide average performance with 90\% bootstrap confidence intervals.  
Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}, on which our agents are also based.\footnote{rllab can be found at \url{https://github.com/rll/rllab}, while the link to our implementation is not reported for blindness but it is already publically available.}
More details on meta-parameters and exhaustive task descriptions are provided in Appendix \ref{app:exp}.

@@

Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on a continuous variant of the classical Cart-pole task, which is a 2D balancing task. {\color{blue}Despite using more trajectories for each parameter update}, our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.

The Swimmer task is a 3D continuous-control locomotion task. This is more difficult than cart-pole. In particular, the longer horizon and the more complex dynamics can have a dangerous impact on the variance of importance weights. In this case, the self-normalization technique proposed in Section \ref{sec:prac} brings an improvement, as shown in Figure \ref{fig:swimmertwo}. Although the confidence intervals are never disjoint, the average of the self-normalized version is always higher, which we consider enough to employ the variant in the next comparison. Figure \ref{fig:swimmerone} shows self-normalized SVRPG against G(PO)MDP. On average, our algorithm dominates for almost the entire task. Also here, we note an increase of speed in early iterations. Toward the end of the task, the improvement becomes statistically significant.

Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for on-line learning since it is related to an average regret~\citep[\eg][]{duan2016benchmarking}. This corresponds to the area below the curves in the figures. Note that the convergence result Theorem \ref{theo:convergence} is precisely about average return. We also report normalized standard deviation and 90\% bootstrap confidence intervals.
In all the tasks, our algorithm has higher average return and lower normalized standard deviation.
In Cart-pole, disjoint confidence intervals show that the advantage is statistically significant.

\textbf{Preliminary results on actor-critic.} %As mentioned, 
Another variance-reduction technique in policy gradient is to employ baselines, or \textit{critics}. This tool is orthogonal to the methods described in this paper, and the theoretical results of Section \ref{sec:conv} are general in this sense. In the experiments described so far, we compared against the so-called \textit{actor-only} G(PO)MDP, \ie without baseline. In order to move towards a more general understanding of the problem of variance in policy gradient, we also test SVRPG in an \textit{actor-critic} scenario. To do so, we use the more challenging MuJoCo~\citep{todorov2012mujoco} Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer. Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP\footnote{\cite{duan2016benchmarking} report results on REINFORCE. However, inspection on \textit{rllab} code and documentation reveals that it is actually PGT \cite{sutton2000policy}, which is equivalent to G(PO)MDP \citep[shown by][]{peters2008reinforcement}. Using the name REINFORCE in a general way is inaccurate, but widespread.} on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms. Results are promising, showing that a combination of baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.

\section{Conclusion}
In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
The control problem considered in the paper has a series of difficulties that are not common in SL.
Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy-dependent.
We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
%{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
%The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
%In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
%}
%\todomatout{Repetition of Section 6?}
Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
Preliminary results support the effectiveness of SVRPG also with a commonly used baseline for the policy gradient.
Despite that, we believe that it will be possible to derive a baseline specifically designed for SVRPG in order to jointly exploit the RL structure and the SVRG idea. Future work should also focus on making batch sizes $N$ and $B$ adaptive, as suggested in \cite{papini2017adaptive}.

>>>>

\section{Experiments}\label{sec:exp}
In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks: Cart-pole balancing and Swimmer~\citep[\eg][]{duan2016benchmarking}.
We consider G(PO)MDP since has a smaller variance than REINFORCE.
% This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases. 
For our algorithm, we use a batch size $N=100$, a mini-batch size $B=10$, and the jointly adaptive step size $\alpha$ and epoch length $m$ proposed in Section \ref{sec:stopping}. Since the aim of this comparison is to show the improvement that SVRG-flavored variance reduction brings to SG in the policy gradient framework, we set the batch size of the baseline policy gradient algorithm to $B$. In this sense, we measure the improvement yielded by computing snapshot gradients and using them to adjust parameter updates. Since we evaluate on-line performance over the number of sampled trajectories, the cost of computing such snapshot gradients is automatically taken into consideration. To make the comparison fair, we also use Adam in the baseline PG algorithm, which we will denote simply as G(PO)MDP in the following.
In all the experiments we used deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}).
Each experiment is run $10$ times with a random policy initialization and seed, but this initialization is shared among the algorithms under comparison.
The length of the experiment, \ie the total number of trajectories, is fixed for each task. Performance is evaluated by using test-trajectories on a subset of the policies considered during the learning process. We provide average performance with 90\% bootstrap confidence intervals.  
Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}, on which our agents are also based.\footnote{rllab can be found at \url{https://github.com/rll/rllab}, while the link to our implementation is not reported for blindness, but it is already publically available.}
More details on meta-parameters and exhaustive task descriptions are provided in Appendix \ref{app:exp}.

@@

Figure \ref{fig:cartpole} compares SVRPG with G(PO)MDP on a continuous variant of the classical Cart-pole task, which is a 2D balancing task. Despite using more trajectories on average for each parameter update, our algorithm shows faster convergence, which can be ascribed to the better quality of updates due to variance reduction.

The Swimmer task is a 3D continuous-control locomotion task. This task is more difficult than cart-pole. In particular, the longer horizon and the more complex dynamics can have a dangerous impact on the variance of importance weights. In this case, the self-normalization technique proposed in Section \ref{sec:prac} brings an improvement (even if not statistically significant), as shown in Figure \ref{fig:swimmertwo}. 
% Although the confidence intervals are never disjoint, the average of the self-normalized version is always higher, which we consider enough to employ the variant in the next comparison. 
Figure \ref{fig:swimmerone} shows self-normalized SVRPG against G(PO)MDP. Our algorithm outperforms G(PO)MDP for almost the entire learning process. Also here, we note an increase of speed in early iterations, and, toward the end of the learning process, the improvement becomes statistically significant.

Table \ref{tab:1} compares G(PO)MDP and SVRPG in terms of the average return computed on the whole learning process, which is a common metric for online learning since it is related to an average regret~\citep[\eg][]{duan2016benchmarking}. This metric corresponds to the area below the curves in the figures. Note that the convergence result in Theorem \ref{theo:convergence} is precisely about the average return. We also report normalized standard deviations and 90\% $t$-confidence intervals.
In both tasks, our algorithm has higher average return and lower normalized standard deviation.
In Cart-pole, disjoint confidence intervals show that the advantage is statistically significant.

\textbf{Preliminary results on actor-critic.} %As mentioned, 
Another variance-reduction technique in policy gradient is to employ baselines or \textit{critics}. This tool is orthogonal to the methods described in this paper, and the theoretical results of Section \ref{sec:conv} are general in this sense. In the experiments described so far, we compared against the so-called \textit{actor-only} G(PO)MDP, \ie without the baseline. To move towards a more general understanding of the problem of variance in policy gradient, we also test SVRPG in an \textit{actor-critic} scenario. To do so, we consider the more challenging MuJoCo~\citep{todorov2012mujoco} Half-cheetah task, a 3D locomotion task that has a larger state-action space than Swimmer. Table \ref{tab:1} also reports the comparison between SVRPG and G(PO)MDP\footnote{\citet{duan2016benchmarking} report results on REINFORCE. However, inspection on \textit{rllab} code and documentation reveals that it is actually PGT \cite{sutton2000policy}, which is equivalent to G(PO)MDP \citep[shown by][]{peters2008reinforcement}. Using the name REINFORCE in a general way is inaccurate, but widespread.} on Half-cheetah using the critic suggested in \cite{duan2016benchmarking} for both algorithms. Results are promising, showing that a combination of the baseline usage and SVRG-like variance reduction can yield an improvement that the two techniques alone are not able to achieve.

\section{Conclusion}
In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
The control problem considered in the paper has a series of difficulties that are not common in SL.
Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy-dependent.
We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
%{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
%The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
%In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
%}
%\todomatout{Repetition of Section 6?}
Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
Preliminary results support the effectiveness of SVRPG also with a commonly used baseline for the policy gradient.
Despite that, we believe that it will be possible to derive a baseline designed explicitly for SVRPG to exploit the RL structure and the SVRG idea jointly. Future work should also focus on making batch sizes $N$ and $B$ adaptive, as suggested in \cite{papini2017adaptive}.
